---
title: "rmet2 Working Example"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

This package is designed to assist members of the general public with running US EPA's Meteorological Processors, [AERMINUTE][https://www3.epa.gov/scram001/metobsdata_procaccprogs.htm], [AERSURFACE][https://www3.epa.gov/ttn/scram/dispersion_related.htm#aersurface], and [AERMET][https://www3.epa.gov/scram001/metobsdata_procaccprogs.htm]. Rmet2 provides the following features:

* Automatic Downloading of Meteorological Data from National Oceonographic and Atmospheric Administration (NOAA) websites.

* Scripted set up of runstream input files to assist in reproducibility.

* Caputuring output runstreams and files and producing tables and figures useful for air dispersion modelers to perform quality assurance checks on intermediate and final outputs of the meteorological preprocessors.


## Example of usage - Corpus Christi, TX

To use rmet2, you will need 

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
