---
title: "README"
output: github_document
always_allow_html: yes


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

This package is designed to assist running US EPA's Meteorological Processors, [AERMINUTE](https://www3.epa.gov/scram001/metobsdata_procaccprogs.htm), [AERSURFACE](https://www3.epa.gov/ttn/scram/dispersion_related.htm#aersurface), and [AERMET](https://www3.epa.gov/scram001/metobsdata_procaccprogs.htm). Rmet2 provides the following features:

* Automatic Downloading of Meteorological Data from National Oceonographic and Atmospheric Administration (NOAA) websites.

* Scripted set up of runstream input files to assist in reproducibility.

* Caputuring output runstreams and files and producing tables and figures useful for air dispersion modelers to perform quality assurance checks on intermediate and final outputs of the meteorological preprocessors.

The package is designed to work with [R](www.r-project.org). 

## Installation
In R, the following script will check for, and if needed, install the package [devtools](https://cran.r-project.org/web/packages/devtools/index.html), and then attempt to install rmet2 to your system. You will only have to install rmet2 once on your system, so for normal day-to-day use you will not need to run the installation script:

```{r installrmet2}
if(!"devtools" %in% installed.packages()) install.pacakges(devtools)
if(!"rmet2" %in% installed.packages()) devtools::install_git("https://github.com/YoJimboDurant/rmet2")
```

TO use rmet2, you will type the following into the console:

```{r libraryrmet2, message=FALSE, warning=FALSE}
library(rmet2)
```


```{r installAM}
installAM()

```
This is the location of the executables.

```{r check}
sapply(c("aermet", "aerminute", "aersurface"), getOption)
```
## Examples usage:

These are updated to work with the new version of aermet using the `createInput(KORH, type = c("aermet23"))`. It will create the combined stage 2 input file.

Additionally, the script has been updated to remove dependencies on packages that are
no longer on CRAN (mainly due to rgdal issues).

[Worcester Regional Airport](https://github.com/YoJimboDurant/rmet2/blob/master/examples/MA_KORH.R)

