---
title: "Corpus Christi Example (KCRP)"
output: github_document
always_allow_html: yes


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Example of usage - Interactive Single Site
For a single site, most users will develop and run a script to process the data. First, we need to identify NOAA surface stations near the site of interest.

TO use rmet2, you will type the following into the console:

```{r libraryrmet2, message=FALSE, warning=FALSE}
library(rmet2)
```

The `mapstation` function can be used to identify prospective stations near a site:
```{r map}
mapstation("Corpus Christi, TX", labels="text")
```

Data for the project:

* We select Corpus Christi International Airport as our surface station (USAF ID=722510, WBAN = 12924).

* We will use upper air data from Corpus Christi (WBAN = 12924, WMO = 72251)

* We will run AERMET for the January 1, 2009 - December, 31, 2013 (5 years)

* We will process using AERMINUTE

* For purposes of demostration, we will run the project in directory "C:/temp/CC"

### Install Aermet, Aerminute, and Aersurface
`installAM` attempts to download the current versions of Aermet, Aersurface and Aerminute from the [SCRAM Meteorological Processors website](https://www3.epa.gov/scram001/metobsdata_procaccprogs.htm) and [SCRAM  Related Programs Website](https://www3.epa.gov/scram001/dispersion_related.htm). After downloading, the program sets the location of aermet, aerminute, and aersurface executables in the `options()`. Links at these websites change, so `installAM` can be set just to configure `options` - see `?installAM`.

```{r installAM}
installAM()

```
This is the location of the executables.

```{r check}
sapply(c("aermet", "aerminute", "aersurface"), getOption)
```


### Create rmet object

```{r include=FALSE}
if(file.exists("C:/Temp/cc/CORPUS.CHRISTI.AP.rds")) file.remove("C:/Temp/cc/CORPUS.CHRISTI.AP.rds")
```
```{r createObj, message = FALSE}
kcrp <- createMetProject(
  project_Name = "CORPUS CHRISTI AP",
   project_Dir="C:/temp/cc",
  start_Date = lubridate::mdy_hm("01/01/2009 00:00", tz="Etc/GMT+6"),
  end_Date = lubridate::mdy_hm("12/31/2013 23:00", tz="Etc/GMT+6"),
  surf_UTC = -6,
  surf_WBAN = 12924,
  surf_USAF = 722510,
  surf_Call="KCRP",
  surf_Latitude = 27.7742,
  surf_Longitude = -97.5122,
  surf_Elevation = 13,
  surf_AnenometerHeight = 10,
  ua_WMO=72251,
  ua_UTC=-6,
  ua_Latitude = 27.77,
  ua_Longitude = -97.50,
  ifg = "Y 02 13 2006",
  lc_File = "C:\\temp\\cc\\landuse\\kcrp_nlcd92.tif",
  lc_Type = "NLCD1992",
  imp_File = NULL,
  imp_Type = NULL,
  cnpy_File = NULL,
  cnpy_Type = NULL,
  as_Snow = "N",
  as_Arid = "Y",
  as_Moisture = "A",
  as_Airport = "Y",
  as_Winter_NS ="12 1 2",
  as_Winter_WS = NULL,
  as_Spring ="3 4 5",
  as_Summer = "6 7 8",
  as_Autumn = "9 10 11")
```

The project is also save as an RDS file. If the RDS file is detected in the directory, `createMetProject` will automatically load that and issue a warning. 

#### Download Data
Now we can progress to downloading data
```{r download, cache=TRUE}
downloadTD3505(kcrp) # could combine into one step
downloadTD6405(kcrp)
downloadTD6401(kcrp)
downloadFSL(kcrp)
```

#### Create Input Files
Two functions are used to create input strings for the AERMET and AERSURFACE and then write them to the appropriate directories
```{r inputam}
kcrp <- createInput(kcrp) # probably should combine this into one step
kcrp <- writeInputFile(kcrp, c("aerminute", "aersurface"))
```


### Process Data
```{r aerminute_surface}
kcrp <- processMet(kcrp, processor=c("aerminute", "aersurface", "aermet1", "aermet2", "aermet3"))
```

### Finalize Main SFC and PFL files
```{r finalize}
makeFinal(kcrp, "kcrp")